---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
## Loading and preprocessing the data
setwd("D:/Reproducible research/RepData_PeerAssessment1")
unzip("./activity.zip")
data <- read.csv("./activity.csv",stringsAsFactors = FALSE, na.strings = NA)
data$date <- as.Date(data$date, "%Y-%m-%d")
Sys.setlocale("LC_TIME","English")
```

Histogram of the total number of steps taken each day

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
library(dplyr)
data <- tbl_df(data)
## What is mean total number of steps taken per day?
stepsxday <- data %>% group_by(date) %>% summarise(total_steps = sum(steps, na.rm = T),
                                      mean_steps = mean(steps, na.rm = T))
library(ggplot2)
ggplot(stepsxday,aes(total_steps))+geom_histogram()+labs(x="total number of steps")
```

Mean and median number of steps taken each day

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
summary(stepsxday[,2:3])
```

Time series plot of the average number of steps taken

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
## What is the average daily activity pattern?
stepsxinterval <- data %>% group_by(interval) %>% 
        summarise(mean_steps = mean(steps, na.rm = T))

ggplot(stepsxinterval, aes(interval,mean_steps))+ geom_line()+ labs(x="5-minute interval", y = " average number of steps taken")
```

The 5-minute interval that, on average, contains the maximum number of steps

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
stepsxinterval$interval[stepsxinterval$mean_steps==max(stepsxinterval$mean_steps)]
```

Code to describe and show a strategy for imputing missing data

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
## Imputing missing values
any(is.na(data[,2:3])) # There is only NAs in the fist column
data1 <- data
for(i in seq_along(data1$steps)) {
        if(is.na(data1[i,1])) {
                row <- stepsxinterval$interval
                data1$steps[i] <- subset(stepsxinterval$mean_steps,row == data$interval[i])     }
}
```

Histogram of the total number of steps taken each day after missing values are imputed

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
stepsxday1 <- data1 %>% group_by(date) %>% summarise(total_steps = sum(steps, na.rm = T),
                                      mean_steps = mean(steps, na.rm = T))

ggplot(stepsxday1,aes(total_steps))+geom_histogram()+labs(x="total number of steps")

summary(stepsxday1[,2:3])
```
                The mean and median of de total number of steps taken each day after missing values are imputed are similar because I use the mean for replace the missing values.
                
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, message = FALSE, warning=FALSE,echo=TRUE } 
## Are there differences in activity patterns between weekdays and weekends?
data1$typeday <- ifelse(weekdays(data1$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")

mean_wy <- data1 %>% filter(typeday == "weekday") %>% group_by(interval) %>%
        summarise(steps= mean(steps)) %>% mutate(typeday = "weekdays")

mean_wd <- data1 %>% filter(typeday == "weekend") %>% group_by(interval) %>%
        summarise(steps= mean(steps)) %>% mutate(typeday = "weekends")

mean_by_type <-bind_rows(mean_wy ,mean_wd)

ggplot(mean_by_type,aes(interval,steps))+geom_line()+ facet_grid(typeday~.)+ labs(x = "5-minute interval", y = "number of steps taken")
```

There are differences in activity patterns between weekdays and weekends, in weekdays there are higher activity before the 10000th 5-minute interval, after these there are higher activity in weekends.